---
title: "Assignment 3 - First Steps"
author:
  - name: Jae-Young Son
  - name: Joey Heffner
  - name: James Wilmott
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The original study

We are replicating portions of the study 'Maps of subjective feelings' by Lauri Nummenmaa, Riitta Hari, Jari K. Hietanen, and Enrico Glerean. The doi for the study is found at: https://www.pnas.org/content/115/37/9198 

In the study, Experiment 1 investigated 


## Goal of assignment

Replicate tables and figures for Experiment 1 (ratings on a variety of dimensions for different subjective feelings)


## Step 0. Clean and import the data

```{r}

path <- dirname(rstudioapi::getSourceEditorContext()$path)

libraryBooks <- c("knitr", "tidyverse", "cowplot")
invisible(lapply(libraryBooks, require, character.only = TRUE)); rm(libraryBooks)

```



## Step 1. Get data for 4 ratings of each emotion (4 basic dimensions are bodily saliency, mental saliency, emotion, and controllability) 

```{r}

```



## Step 2. Plot the histograms from Figure 1 for Experiment 1


```{r}

```



## Step 3. Run Pearson correlations without clustering to get an idea of how these dimensions of feeling are related to each other

```{r}

```



```{r}
'End of document'
```

